---
title: "Tagging CTD Data"
author: "Dan Kelley (https://orcid.org/0000-0001-7808-5911)"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    toc: true
    number_sections: true
    fig_caption: yes
    fig_width: 3.5
    fig_height: 3.5
    dpi: 72
    dev.args: list(pointsize=11)
vignette: >
  %\VignetteIndexEntry{Tagging CTD Data}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options:
  markdown:
    wrap: 72
---

```{r, echo = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

**Abstract.** This vignette shows how to use low-level `ocetag` functions to
set up a tagging database, how to add entries to it, and how to recover
entries.

First, read a CTD file and plot a density profile. You may ignore the warnings,
one of which indicates that the file has an erroneous start-time, the other of
which indicates that the file provides temperature on an archaic scale, which
gets auto-converted by `read.oce()`.

In the plot, you may notice evidence of a mixed layer within the top 10m of the
water column. The code block below is written so that `locator()` will be
called if you run it interactively and otherwise, as in the building of this
vignette, the result is inserted directly. Note that the `ctdtag()` function
provides better ways to indicate such points, but the purpose here is to show
how to use the low-level functions.

```{r}
library(ocetag)
library(oce)
name <- "ctd.cnv"
f <- system.file("extdata", name, package = "ocetag")
d <- oce::read.oce(f)
plotProfile(d, xtype = "sigma0", type = "p", cex = 0.3)
# Change next to TRUE for interactive work
if (FALSE) {
    xy <- locator(1)
    usr <- par("usr")
    dx <- (xy$x - d[["sigma0"]]) / (usr[2] - usr[1])
    dy <- (xy$y - d[["pressure"]]) / (usr[4] - usr[3])
    index <- which.min(dx^2 + dy^2)
} else {
    index <- 12
}
points(d[["sigma0"]][index], d[["pressure"]][index], pch = 3, col = 2, cex = 2)
mtext(sprintf("Red cross at index %d", index), side = 1, col = 2, line = -1)
```

Now, we know the point to tag, and the next step is to tag it.  First, we
create a database.  (To satisfy rules of CRAN packages, here we are creating a
temporary file and then removing it. In actual use, of course we would name a
particular file, and we would not erase it!)

```{r}
# Define the list of tags, here just one, for mixed-layer depth.
mapping <- list("MLD" = 1)
```

```{r}
# Add 'pressure' to the columns in the default 'tags' table.
tags <- list("pressure" = "FLOAT")
```

We now have what we need to for the database work.  First, we create the
database.  (NB. if `createDatabase()` is called again with the same database
name, it will leave the file as it is.)

```{r echo=FALSE}
dbname <- tempfile() # Do not use a tempfile() in actual work!!
```
```{r eval=FALSE}
dbname <- "tags.db"
```
```{r}
createDatabase(dbname, mapping = mapping, tags = tags)
```

Saving the tag is simple. Note that the *numeric* value of the tag is
saved, *not* the string value.
```{r}
saveTag(name,
    index = index, pressure = d[["pressure"]][index],
    tag = 1, analyst = "Dan Kelley", dbname = dbname
)
```

Reading the tags later is also simple.
```{r}
print(getTags(name, dbname = dbname))
```

For readers who are familiar with SQL syntax, note that typing
```
echo .dump | sqlite3 dbname
```
in a unix terminal yields as follows.
```{sql eval=FALSE}
PRAGMA foreign_keys=OFF;
BEGIN TRANSACTION;
CREATE TABLE `version` (
  `version` INTEGER
);
INSERT INTO version VALUES(1);
CREATE TABLE `mapping` (
  `value` INTEGER,
  `name` TEXT
);
INSERT INTO mapping VALUES(1,'MLD');
CREATE TABLE `tags` (
  `file` TEXT,
  `index` INT,
  `pressure` FLOAT,
  `tag` INT,
  `analyst` TEXT,
  `analysisTime` TIMESTAMP
);
INSERT INTO tags VALUES('ctd.cnv',12,4.197000000000000063,1,'Dan Kelley',1713631216);
COMMIT;
```

```{r echo=FALSE}
unlink(dbname) # Do not erase your efforts in actual work!!
```

